---
title: "gtsummary in quarto"
format:
  html:
    theme: lux
execute:
  echo: false
---

```{r}
#| message: false
#dplyr installed
library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd",
               "id", "nsibs", "samp", "race_eth", "sex", "region",
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"),
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |>
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes")))
```

The **NLSY79 Cohort is a longitudinal project that follows the lives of a sample of American youth born between 1957-1964.** The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979; after two subsamples were dropped, 9,964 respondents remain in the eligible samples. Data are now available from Round 1 (1979 survey year) to Round 30 (2022 survey year).

- bullet

- Number

- Hello\
  1.

  2.  

# Customization of `tbl_summary()`

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

```{r}
tbl_summary(
 nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

```{r}
#| label: tbl-sleep
#| tbl-cap: "Descriptive statistics including sleep"
sleep_table <- tbl_summary(nlsy,
						include = c(starts_with("sleep"), sex_cat),
						by = region_cat)|>
	add_overall()
sleep_table
```
@tbl-one shows the descriptive statistics about sleep

```{r}
sleep_median <-median(nlsy$sleep_wknd, na.rm = TRUE)
sleep_median
```
The median number hours of sleep during the weekend was `r sleep_median`

```{r}
sleep_table$table_body
inline_text(sleep_table, variable = "sleep_wknd",
						column = "stat_0")

inline_text(sleep_table, variable = "sex_cat",
 						level = "Female",
 						column = "stat_0")
inline_text (sleep_table, variable = "sleep_wknd",
            pattern = "{median}, IQR= ({p25}, {p75})")
```

`In the sample the average sleep on weekends was `r inline_text(sleep_table, variable = "sleep_wknd", column = "stat_0")`

```{r}
median_iqr <- inline_text(sleep_table, variable = "sleep_wknd",
						column = "stat_0")
just_median <- inline_text(sleep_table, variable = "sleep_wknd",
						column = "stat_0", pattern = "{median}")
median_text <- inline_text(sleep_table, variable = "sleep_wknd",
						column = "stat_0", pattern = "{median}, IQR = [{p25}, {p75}]")
```


There are multiple ways we could report the median, `r median_iqr`, or we could say `r just_median`, or finally `r median_text`.

```{r}
#| label: tbl-sleep-means
#| tbl-cap: "Descriptive statistics including sleep"
sleep_table2 <- tbl_summary(nlsy,
						include = c(starts_with("sleep"), sex_cat),
						by = region_cat,
						statistic = list(all_continuous() ~ "{mean}, sd = {sd}")) |>
	add_overall()
sleep_table2
```

```{r}
inline_text(sleep_table2, variable = "sleep_wknd",
						column = "stat_0")

inline_text(sleep_table2, variable = "sleep_wknd",
						column = "stat_0", pattern = "{mean}")
```
